---
title: "FINAL PROJECT (KMEANS)"
author: "Ummar A. Sallil"
date: "2022-12-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Helper packages
```{r}
library(dplyr)      
library(ggplot2)     
library(stringr)     
library(gridExtra)   
```
# Modeling packages
```{r}
library(tidyverse)  
library(cluster)     
library(factoextra)  
```
# loading dataset
```{r}
radiomics <- read.csv("radiomics.csv")
attach(radiomics)
```
```{r}
X1 <- radiomics %>% 
  select_if(is.numeric)
```
#we start by scaling/standardizing the data

```{r}
df <- scale(X1)
head(df)
```
# Identifying cluster with different centers
```{r}
cluster0 <- kmeans(df, centers = 2, nstart = 25)
cluster1 <- kmeans(df, centers = 3, nstart = 25)
str(cluster1)
cluster2 <- kmeans(df, centers = 4, nstart = 25)
str(cluster2)
cluster3 <- kmeans(df, centers = 5, nstart = 25)
str(cluster3)
```
#plotting the clustered results above
```{r}
fviz_cluster(cluster0, data = df)
fviz_cluster(cluster1, data = df)
fviz_cluster(cluster2, data = df)
fviz_cluster(cluster3, data = df)
```
# based on the 3 results, cluster0 seems to be a good clustering since no overlapping of the entries.  

# Creating a plot of the number of cluster using nbclust with method within sum of squares and clusGap.
```{r}
fviz_nbclust(df, kmeans, method = "wss")
```

```{r}
gap_stat <- clusGap(df,
                    FUN = kmeans,
                    nstart = 25,
                    K.max = 10,
                    B = 25)
fviz_gap_stat(gap_stat)
```
# The results of the two approaches above confirm that the optimum cluster is equal to 2.

# dataset reproducible
```{r}
set.seed(143)
```

# adding clust0 to the original dataset to have the final data. 
```{r}
final_data <- cbind(radiomics, cluster = cluster0$cluster)
head(final_data)
```




