---
title: "Final Project (Hierarchical Clustering)"
author: "Ummar A. Sallil"
date: "2022-12-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Helper Packages
```{r}
library(dplyr)       # for data manipulation
library(ggplot2)     # for data visualization
library(cluster)     # for general clustering algorithms
library(factoextra)  # for visualizing cluster results
library(readr)
library(modeldata)
library(dplyr) 
library(tidyverse)
library(recipes)
```
# loading dataset
```{r}
radiomics <- read.csv("radiomics.csv")
attach(radiomics)
```
```{r}
X1 <- radiomics %>% 
  select_if(is.numeric)
```
#we start by scaling/standardizing the data

```{r}
df <- scale(X1)
head(df)
```
# Identifying best method to be used.
```{r}
method <- c( "average", "single", "complete", "ward")
names(method) <- c( "average", "single", "complete", "ward")
```
```{r}
K<- function(x) {
  agnes(df, method = x)$ac
}

sapply(method, K)
```
# the result shows that ward has a lowest variance sin it produces the highest agglomerative coefficient.
```{r}
clust <- agnes(df, method = "ward")
```
# creating dendogram tree
```{r}
dendotree <-pltree(clust, cex = 0.6, hang = -1, main = "Dendrogram")
```
#calculate gap statistic for each number of clusters (up to 10 clusters)
```{r}
gap_stat <- clusGap(df, FUN = hcut, nstart = 25, K.max = 10,)
```
#produce plot of clusters vs. gap statistic

```{r}
fviz_gap_stat(gap_stat)
```
#compute distance matrix

```{r}
d <- dist(df, method = "euclidean")
```
#perform hierarchical clustering using Ward's method

```{r}
final_clust <- hclust(d, method = "ward.D2" )
```
#cutting the dendrogram into 2 clusters

```{r}
dendonew <- cutree(final_clust, k = 2)
```
#find number of observations in each cluster
```{r}
table(dendonew)
```
#append cluster labels to original data
```{r}
final_data <- cbind(radiomics, cluster = dendonew)
head(final_data)

```
# Plotting cluster with different methods for comparison
```{r}
p1 <- fviz_nbclust(df, FUN = hcut, method = "wss", 
                   k.max = 10) +
  ggtitle("(A) Elbow method")
p2 <- fviz_nbclust(df, FUN = hcut, method = "silhouette", 
                   k.max = 10) +
  ggtitle("(B) Silhouette method")
p3 <- fviz_nbclust(df, FUN = hcut, method = "gap_stat", 
                   k.max = 10) +
  ggtitle("(C) Gap statistic")
```
# Display plots side by side
```{r}
gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
```
#Plot full dendogram
```{r}
agn  <- as.dendrogram(as.hclust(clust))
dagn2 <- as.dendrogram(as.hclust(clust), hang = 0.2)
op <- par(mar = par("mar") + c(0,0,0, 2)) # more space to the right
plot(dagn2, horiz = TRUE)
plot(agn, horiz = TRUE, center = TRUE,
     nodePar = list(lab.cex = 0.6, lab.col = "forest green", pch = NA),
     main = deparse(agn$call))
par(op)

```

###------------------ END------------------###







