---
title: "STT371 FINAL PROJECT (Model 1)"
author: "Ummar A. Sallil"
date: "2022-12-12"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
``````
# Helper Packages
```{r}
library(ggplot2)
library(lattice)
library(caret)
library(readr)
library(rsample)
library(modeldata)
library(dplyr) 
library(tidyverse)
library(ROCR)
library(pROC)
library(xgboost)
library(recipes)
```
# Loading dataset
```{r}
radiomics<- read.csv("radiomics.csv")
```

```{r}
head(radiomics)
```

```{r}
summary(radiomics)
```

## Processing the dataset
#Check for null and missing values
```{r}
sum(is.na(radiomics))
```
# Checking for normality of the dataset using shapiro-test
```{r}
library(dplyr)
X1 <- radiomics %>% 
  select_if(is.numeric)
```
#If the p-value of the test is greater than α = .05, then the data is assumed to be normally distributed.

```{r}
X2 <- apply(X1,2,shapiro.test)
```
# since all variable were tested, we need to determine which variables are skewed. 
```{r}
normaldata <- unlist(lapply(X2, function(x) x$p.value))
sum(normaldata>0,05)
skew <- names(normaldata[normaldata>0.05])
skew
```
#the result shows that an "Entropy_cooc.W.ADC" is skewed or not normally distributed.
# normalizing the skewed data
```{r}
Nor <- log(radiomics$Entropy_cooc.W.ADC)
shapiro.test(Nor)
```
# Get the correlation of the whole dataset expect the categorical variables
```{r}
library(dplyr)
newdf1 = select(radiomics, -c("Institution","Failure.binary", "Failure"))
cor.newdf1 = cor(newdf1)
corr = round(cor.newdf1,2) 
head(corr)
```
# Data preparation and splitting using only the normal dataset
```{r}
set.seed(123)  # for reproducibility
df <- radiomics %>% mutate_if(is.ordered, factor, ordered = FALSE)
df <- radiomics %>%
  mutate_if(str_detect(names(.), 'Qual|Cond|QC|Qu'), as.numeric)

```
#Split the data intro training (80%) and testing (20%) stratified in Failure.binary column

```{r}
split = initial_split(df,prop = 0.8 ,strata = "Failure.binary")
churn_train <- training(split)
churn_test  <- testing(split)

```
# pre-processing for X and Y using xgboost
```{r}
xgb_train <- recipe(Failure.binary~ ., data = df) %>%
  step_integer(all_nominal()) %>%
  step_nzv(all_nominal()) %>%
  step_integer(contains("Dissimilarity")) %>%
  step_integer(Failure) %>%
  step_integer(Entropy_cooc.W.ADC) %>%
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes()) %>%
  prep(training = churn_train, retain = TRUE) %>%
  juice()
```
```{r}
X <- as.matrix(xgb_train[setdiff(names(xgb_train), "Failure.binary")])
Y <- xgb_train$Failure.binary
```
# setting optimal parameter list of the dataset
```{r}
params <- list(
  eta = 0.01,
  max_depth = 3,
  min_child_weight = 3,
  subsample = 0.5,
  colsample_bytree = 0.5
)
```

```{r}
xgb.fit.final <- xgboost(
  params = params,
  data = X,
  label = Y,
  nrounds = 999,
  objective = "binary:logistic",
  verbose = 0
)
```
```{r}
summary(xgb.fit.final)
```
#Print the Top 20 important features during Training

```{r}
vip::vip(xgb.fit.final, num_features = 20)
```
# Prediction performance of the model using training dataset
```{r}
pred1<- predict(xgb.fit.final, X, type = "prob")
pred1
perf1 <- prediction(pred1, churn_train$Failure.binary) %>%
  performance(measure = "tpr", x.measure = "fpr")

```
```{r}
par(mfrow = c(1,2))
```
# Training model prediction performance
```{r}
roc(churn_train$Failure.binary ~ pred1, 
    plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, 
    col="red", 
    lwd=2, 
    print.auc=TRUE, 
    main = "Performance in Training")
```
###__________________________________#####__________________________________####

#Print the AUC values during Testing
```{r}
xgb_test <- recipe(Failure.binary~ ., data = churn_test) %>%
  step_integer(all_nominal()) %>%
  step_nzv(all_nominal()) %>%
  step_integer(contains("Dissimilarity")) %>%
  step_integer(Failure) %>%
  step_integer(Entropy_cooc.W.ADC) %>%
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes()) %>%
  prep(training = churn_test, retain = TRUE) %>%
  juice()
```

```{r}
X1 <- as.matrix(xgb_test[setdiff(names(xgb_test), "Failure.binary")])
```
```{r}
pred_test<- predict(xgb.fit.final, X1, type = "prob")
pred_test
perf2 <- prediction(pred_test, churn_test$Failure.binary) %>%
  performance(measure = "tpr", x.measure = "fpr")
```
# Testing set prediction performance
```{r}
roc(churn_test$Failure.binary ~ pred_test, plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, col="red", lwd=2, print.auc=TRUE, main = "Performance in testing")

```

#################### END #############